### 2015-03-12 article/article.tex
commit

Since annotation from multiple public databases and internal data were used, there was a need to design and implement a database supporting the research. The database structure thereby presents 7 tables and aims to facilitate the processes of gathering, parsing and statistical analysis of the nuclear receptors data, as well as to allow for a clear mapping between the nuclear receptor genes and the phenotypes and SNPs associated with them. 
The main challenge was to establish a connection between the MGI, MPD and UCSC tables. Therefore, there are 2 tables containing information from the MGI, 3 tables for MPD, 1 USCS table and the core table:
\begin{itemize}
\item \textbf{nr\_mapping} - core table, which contains the associations between the 49 nuclear receptors in the mouse and their corresponding gene names. 
\item \textbf{mgi} - contains the MGI gene annotations (MGI gene id, type, attributes, transmission etc.). 
\item \textbf{mgi\_phenotypes} - stores phenotypic information in association with MGI genes; more than one phenotype can be associated with a specific gene.
\item \textbf{mpd\_snp} - contains the MPD SNP-annotations in correlation with the nuclear receptor genes. 
\item \textbf{ucsc} - contains a more detailed overview on the MPD mutations from \textit{mpd\_snp} 
\item \textbf{mpd\_strains} - contains the MPD strain annotations (MPD strain id, sex, number of mice, Z-score etc.)
\item \textbf{mpd\_phenotypes} - stores phenotypic information in association with MPD strains; more than one phenotype can be associated with a specific strain.
\end{itemize} 
~~~~~~~\\  
\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/db.png}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{Database}. 
	~~~~~~~\\
	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
	\label{fig:mgi_pheotypes_distribution}
\end{figure}
%------------------------------------------------

@@

>>>>

Since annotation from multiple public databases and internal data were used, there was a need to design and implement a database supporting the research. The database structure thereby presents 7 tables and aims to facilitate the processes of gathering, parsing and statistical analysis of the nuclear receptors data, as well as to allow for a clear mapping between the nuclear receptor genes and the phenotypes and SNPs associated with them. 
The main challenge was to establish a connection between the MGI, MPD and UCSC tables. Therefore, there are 2 tables containing information from the MGI, 3 tables for MPD, 1 USCS table and the core table:
\begin{itemize}
\item \textbf{nr\_mapping} - core table, which contains the associations between the 49 nuclear receptors in the mouse and their corresponding gene names. 
\item \textbf{mgi} - contains the MGI gene annotations (MGI gene id, type, attributes, transmission etc.). 
\item \textbf{mgi\_phenotypes} - stores phenotypic information in association with MGI genes; more than one phenotype can be associated with a specific gene.
\item \textbf{mpd\_snp} - contains the MPD SNP-annotations in correlation with the nuclear receptor genes. 
\item \textbf{ucsc} - contains a more detailed overview on the MPD mutations from \textit{mpd\_snp} 
\item \textbf{mpd\_strains} - contains the MPD strain annotations (MPD strain id, sex, number of mice, Z-score etc.)
\item \textbf{mpd\_phenotypes} - stores phenotypic information in association with MPD strains; more than one phenotype can be associated with a specific strain.
\end{itemize} 
~~~~~~~\\  
\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/db.png}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{Database}. 
	~~~~~~~\\
	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
	\label{fig:database}
\end{figure}
%------------------------------------------------

@@

%------------------------------------------------
\section{The 49 nuclear receptors in the human genome}

In contrast to the nuclear receptors in mouse, there have been two studies about nuclear receptors in human, called \textit{kora} and \textit{twinsuk}. With the results of these studies the database used in this paper was expanded beyond one species with the goal to find similarities between the nuclear receptors in human and mouse. Because of the lack of researched nuclear receptors in mouse there can be only made comparisons regarding the phenotype with the tables \textit{mgi\_phenotypes} \ref{fig:database} and \textit{mpg\_phenotypes} \ref{fig:database}. Table \ref{tab:compHuMo} shows a list with nuclear receptors which are found in mouse and human and have a similar phenotype or a similar biological function.
~~~~~~~\\
\begin{table}[H]
\centering
\begin{tabulary}{\linewidth}{ LLLLL }
\hline
	\textbf{Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
	\\  
	\hline \hline 
	Nr2e1  & corticosterone, \textit{steroid hormone} & lathosterol, \textit{steroid hormone}\\
	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
	Thrb  & ,\textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
	Vdr & \textit{Osteoporosis} & \textit{Osteoporosis}\\
	Ar & Spinal and bulbar muscular atrophy & Spinal and bulbar muscular atrophy\\
	\hline
\end{tabulary}
\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
~~~~~~~\\
Nuclear receptors in human and mouse which share a similar protein or a similar biological function.}} 
\label{tab:compHuMo}
\end{table}
